Galactic Chemical Evolution Models Favor an Extended Type Ia Supernova Delay-Time Distribution

### 2024-04-02 src/tex/ms.tex
[showyourwork] overleaf sync

% user-defined commands
\newcommand{\yes}{\textcolor{green}{\checkmark}\xspace}
\newcommand{\meh}{\textcolor{black}{$\sim$}\xspace}
\newcommand{\no}{\textcolor{red}{$\times$}\xspace}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}
\newcommand{\ccappaffil}{Center for Cosmology and AstroParticle Physics, The Ohio State University, 191 W. Woodruff Ave., Columbus OH 43210, USA}
\newcommand{\aFe}{[$\alpha$/Fe]\xspace}
\newcommand{\vice}{{\tt VICE}\xspace}
\newcommand{\todo}[1]{{\color{red}#1}}

@@

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation, and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in %one-zone and 
    multi-zone chemical evolution models which include radial stellar migration.
    % using the Versatile Integrator for Chemical Evolution (\vice). \vice includes a prescription for radial stellar migration in multi-zone models. 
    We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the final data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio.
    % whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe.
    While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor an extended DTD with fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

We use \vice to run chemical evolution models which closely follow those of \citet{JohnsonWeinberg2020-Starbursts} and \citet[][hereafter \citetalias{Johnson2021-Migration}]{Johnson2021-Migration}. We refer the interested reader to the former for details about the \vice package and to the latter for details about the model Milky Way disk, including the star formation law, radial density gradient, and outflows. Similar to \citetalias{Johnson2021-Migration}, we adopt a prescription for radial migration based on the {\tt h277} hydrodynamical simulation \citep{Christensen2012-h277}. In Appendix \ref{app:migration}, we describe our method for determining the migration distance $\Delta R_{\rm gal}$ and midplane distance $|z|$ for each model stellar population. Our method produces smoother distributions in chemical abundance space than the simulation-based approach, but the abundance distributions are otherwise unaffected by this change. Table \ref{tab:multizone-parameters} summarizes our model parameters and the sub-sections in which we discuss them in detail.

@@

\paragraph{Early-burst} An extension of the model proposed by \citet{Conroy2022-ThickDisk} to explain the non-monotonic behavior of the high-$\alpha$ sequence down to ${\rm [Fe/H]}\approx-2.5$. This model features an abrupt factor $\sim20$ rise in the SFE at early times, driving an increase in the [O/Fe] abundance at the transition between the epochs of halo and thick disk formation. \citet{Stahlholdt2022-StarFormationEpochs} found evidence for a burst $\sim10$ Gyr ago which marks the beginning of a second phase of star formation. An early infall-driven burst of star formation has been found to lead to MW-like $\alpha$-bimodality in simulations \citep{Mackereth2018-AlphaBimodality}. We adopt the following formula for the time-dependence of the SFE timescale from \citet{Conroy2022-ThickDisk}:
\begin{equation}
    \frac{\tau_{\rm EB}}{1\,\rm{Gyr}} =
    \begin{cases}
        50, & t < 2.5\,\rm{Gyr} \\
        \frac{50}{[1+3(t-2.5)]^2}, & 2.5\leq t \leq 3.7\,\rm{Gyr} \\
        2.36, & t > 3.7\,\rm{Gyr}.
    \end{cases}
    \label{eq:earlyburst-taustar}
\end{equation}
While \citet{Conroy2022-ThickDisk} used a constant infall rate in their one-zone model, we adopt a radially-dependent infall rate which declines exponentially with time:
\begin{equation}
    f_{\rm EB}(t|R_{\rm gal}) = e^{-t/\tau_{\rm sfh}},
    \label{eq:earlyburst-ifr}
\end{equation}
where $\tau_{\rm sfh}$ is the same as in the inside-out case.
To calculate $\dot \Sigma_\star$ from the above quantities, we modify the fiducial star formation law adopted from \citetalias{Johnson2021-Migration}, substituting $\tau_{\rm EB}$ for the SFE timescale of molecular gas:
\begin{equation}
    \dot \Sigma_\star = 
    \begin{cases}
        \Sigma_g \tau_{\rm EB}^{-1}, & \Sigma_g \geq \Sigma_{g,2} \\
        \Sigma_g \tau_{\rm EB}^{-1} \Big(\frac{\Sigma_g}{\Sigma_{g,2}}\Big)^{2.6}, & \Sigma_{g,1} \leq \Sigma_g \leq \Sigma_{g,2} \\
        \Sigma_g \tau_{\rm EB}^{-1} \Big(\frac{\Sigma_{g,1}}{\Sigma_{g,2}}\Big)^{2.6} \Big(\frac{\Sigma_g}{\Sigma_{g,1}}\Big)^{0.7}, & \Sigma_g \leq \Sigma_{g,1},
    \end{cases}
    \label{eq:earlyburst-sfr}
\end{equation}
with $\Sigma_{g,1}=5\times 10^6\,{\rm M}_\odot\,{\rm kpc}^{-2}$ and $\Sigma_{g,2}=2\times 10^7\,{\rm M}_\odot\,{\rm kpc}^{-2}$.

@@

We restrict our sample to red giant branch and red clump stars with high-quality spectra. Table \ref{tab:sample} lists our selection criteria, which largely follow from \citet{Hayden2015-ChemicalCartography}. This produces a final sample of \variable{output/sample_size.txt}stars with [O/Fe] and [Fe/H] abundance measurements. APOGEE stars are matched with their Bailer-Jones photo-geometric distance estimate from \textit{Gaia} Early Data Release 3 \citep{Gaia2016-Mission,Gaia2021-EDR3}, which we use to calculate galactocentric radius $R_{\rm gal}$ and midplane distance $z$.
We use estimated ages from \citet[][hereafter \citetalias{Leung2023-Ages}]{Leung2023-Ages}, who use a variational encoder-decoder network which is trained on asteroseismic data to retrieve age estimates for APOGEE giants without contamination from age-abundance correlations. Importantly, the \citetalias{Leung2023-Ages} ages do not plateau beyond $\sim10$ Gyr as they do in astroNN \citep{Mackereth2019-astroNN-Ages}. We use an age uncertainty cut of 40\% per the recommendations of \citetalias{Leung2023-Ages}, which produces a total sample of \variable{output/age_sample_size.txt}APOGEE stars with age estimates. We note that we use the full sample of \variable{output/sample_size.txt}APOGEE stars unless we explicitly compare to age estimates. Table \ref{tab:sample} also presents the median uncertainties in the data for [Fe/H], [O/Fe], and age.

@@

To quantify the agreement between the MDFs generated by \vice and those observed in APOGEE, we compute the Kullback-Leibler (KL) divergence \citep{KullbackLeibler1951}, defined as
\begin{equation}
    D_{\rm{KL}}(P,Q) \equiv \int_{-\infty}^{\infty} p(x) \log\Big(\frac{p(x)}{q(x)}\Big) dx
    \label{eq:kl-divergence}
\end{equation}
for distributions $P$ and $Q$ with probability density functions $p(x)$ and $q(x)$. If $D_{\rm KL}=0$, the two distributions contain equal information. In this case, $P$ is the APOGEE MDF, $Q$ is the model MDF, and $x={\rm [Fe/H]}$. We bin the distributions with a width of 0.01 dex and apply observational uncertainties to our modeled MDFs according to Table \ref{tab:sample}. For each SFH and DTD, we compute $D_{\rm{KL}}$ in the 18 different Galactocentric regions shown in Figure \ref{fig:feh-df-comparison}. We use bins in $R_{\rm gal}$ with a width of 2 kpc between 3 and 15 kpc, and bins in midplane distance of $|z|=0-0.5$ kpc, $0.5-1$ kpc, and $1-2$ kpc. The score $S$ for the entire model is taken to be the average of $D_{\rm{KL}}$ for each region $R$ weighted by the number of APOGEE stars in that region $N_{\star,R}$:
\begin{equation}
    S = \frac{\sum_R D_{\rm KL}(P,Q|R) N_\star(R)}{\sum_R N_\star(R)}.
    \label{eq:feh-df-score}
\end{equation}

@@

% [Talk about scores]
We again compute the KL divergence (Equation \ref{eq:kl-divergence}) to quantify the agreement between the [O/Fe] DFs of our models and APOGEE. We calculate a score for each model as described in Section \ref{sec:feh-df}. The best-scoring model uses the two-infall SFH with the plateau ($W=0.3$ Gyr) DTD, although both plateau and exponential DTDs as well as the triple-system DTD score well when combined with certain SFHs. The late-burst SFH generally produces [O/Fe] DFs that score poorly compared to other models. [Elaborate a bit more].

@@

% Results for the two-population and power-law ($\alpha=-1.1$) DTDs, with a large fraction of prompt SNe Ia, are presented in the first and second columns, respectively. 
The two-population and power-law ($\alpha=-1.1$) DTDs, which have a large fraction of prompt SNe Ia, produce stellar abundance distributions that are reasonably well-aligned with the APOGEE contours at low $|z|$, but they entirely miss the observed high-$\alpha$ sequence at large $|z|$. The ISM abundance tracks for the 8 kpc zone do not pass through the APOGEE 30\% contour. For both DTDs, the high-[O/Fe] knee is located below the left-most bound of the plot, but we observe a second knee at ${\rm [O/Fe]}\approx0.2$ where the abundance track turns further downward.
% For these models, the simulated stellar abundances are reasonably well-aligned with the APOGEE contours at low $|z|$ (bottom panels), but they entirely miss the observed high-$\alpha$ sequence at high $|z|$ (top panels).

The exponential ($\tau=1.5$ Gyr) DTD produces a distribution in Figure \ref{fig:ofe-feh-dtd} which aligns quite well with the outer APOGEE contours in all $|z|$-bins, and even produces a ``ridge'' which extends to high [O/Fe] at low- and mid-latitudes (bottom and center panels, respectively). While it does better at populating the high-$\alpha$ sequence than the previous DTDs, the bulk of the simulated populations at large $|z|$ still fall below the APOGEE 30\% contour (top panel).

The two right-hand columns present model results for the plateau ($W=1$ Gyr) and triple-system DTDs, which have the longest median delay times. The high-$\alpha$ ``knee'' occurs at a much higher metallicity in these models and is visible in the gas abundance tracks in the upper-left corner of the panels. At high $|z|$, the simulated abundance align quite well with the APOGEE high-$\alpha$ sequence (top panels), but there is a significant ridge of high-$\alpha$ stars from the inner Galaxy at low $|z|$ (bottom panels).

To quantify the agreement between the multi-zone model outputs and data in [O/Fe]--[Fe/H] space, we implement the method of \citet{PerezCruz2008-KLTest2D} for computing the KL divergence between two continuous, multivariate samples using a $k$-nearest neighbor estimate.
% \citet{PerezCruz2008-KLTest2D} present a method for estimating the KL divergence between two continuous, multivariate samples using a $k$-nearest neighbor estimate. We implement this method to calculate the divergence between the multi-zone models and APOGEE data in [O/Fe]--[Fe/H] space. 
\todo{More detail on the scoring calculations.}
% \footnote{We got the implementation from \url{https://mail.python.org/pipermail/scipy-user/2011-May/029521.html}.}

@@

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/radial_migration.pdf}
    \caption{The distribution of final radius $R_{\rm final}$ as a function of formation radius $R_{\rm form}$ and age for the h277 analogue (top row) and Gaussian sampling scheme (bottom row). From left to right, star particles are binned by formation annulus, from the inner disk (far left column) to the solar annulus (center column) to the outer disk (far right column). Within each panel, colored curves represent the different age bins, ranging from the youngest stars (dark blue) to the oldest (dark red). In the top row, we exclude age bins with fewer than 100 unique analogue IDs for clarity. All distributions are normalized so that the area under the curve is 1, and have been smoothed by a boxcar function with a width of 0.5 kpc. The vertical dotted black lines indicate the bounds of each bin in $R_{\rm form}$; stars within that region of the distribution have not migrated outside their birth annulus over their lifetime.}
    \label{fig:radial-migration}
    \script{radial_migration.py}
\end{figure*}

@@

For each star particle in \vice, \citetalias{Johnson2021-Migration} randomly assign an analogue star particle from \texttt{h277} and adopt its radial migration distance $\Delta R$ and final midplane distance $z$. This allows \vice to adopt a realistic pattern of radial migration without needing to implement its own hydrodynamical simulation. However, in regions where the number of h277 star particles is relatively low, such as at large $R_{\rm gal}$ and small $t$, a single h277 star particle can be assigned as an analogue to multiple \vice stellar populations. These populations will have similar formation and migration histories and consequently similar abundances, which produces unphysical ``clumps'' of stars in the abundance distributions at high latitudes and large radii.

We fit a Gaussian to the distribution of $\Delta R = R_{\rm final} - R_{\rm initial}$ from the \texttt{h277} output in bins of $R_{\rm initial}$ and age. Each Gaussian is centered at 0 and we find that the scale $\sigma_{\Delta R}$ is best described by the function
\begin{equation}
    \sigma_{\Delta R} = 1.35\,{\rm kpc} \Big(\frac{R_{\rm form}}{8\,{\rm kpc}}\Big)^{0.61} \Big(\frac{T}{1\,{\rm Gyr}}\Big)^{0.33}
    \label{eq:radial-migration}
\end{equation}
where $T$ is the age of the star particle.

We fit a sech$^2$ function \citep{Spitzer1942} to the distribution of midplane distances $z$. Vertical migration away from the midplane does not affect the chemical evolution model, but we do use $z$ in our analysis. The probability density function (PDF) of $z$ given some scale height $h_z$ is
\begin{equation}
    {\rm PDF}(z) = \frac{1}{4 h_z} {\rm sech}^2\Big(\frac{z}{2 h_z}\Big)
    \label{eq:sech-pdf}
\end{equation}
and the corresponding cumulative distribution function (CDF) is
\begin{equation}
    {\rm CDF}(z) = \frac{1}{1 + e^{-z / h_z}}.
    \label{eq:sech-cdf}
\end{equation}
We fit the above function to the distributions of $z$ in h277 in varying bins of $T$ and $R_{\rm final}$ and found that $h_z$ is best described by the function
\begin{equation}
    h_z = (0.25\,{\rm kpc}) 
    e^{\frac{T-5\,{\rm Gyr}}{7.0\,{\rm Gyr}}}
    e^{\frac{R_{\rm final}-8\,{\rm kpc}}{6.0\,{\rm kpc}}}.
    \label{eq:scale-height}
\end{equation}

@@

\citet{Okalidis2022-AurigaMigration} studied the stellar migration in the Auriga simulations and found that the migration strength had both a dependence on age and formation radius. They also found that in strongly-barred galaxies, the migration is stronger overall but has a slower timestep evolution than diffusion ($\Delta t^{0.5}$).

@@

>>>>

% user-defined commands
\newcommand{\yes}{\textcolor{green}{\checkmark}\xspace}
\newcommand{\meh}{\textcolor{black}{$\sim$}\xspace}
\newcommand{\no}{\textcolor{red}{$\times$}\xspace}
\newcommand{\osuaffil}{Department of Astronomy, The Ohio State University, 140 W. 18th Ave, Columbus OH 43210, USA}
\newcommand{\ccappaffil}{Center for Cosmology and AstroParticle Physics, The Ohio State University, 191 W. Woodruff Ave., Columbus OH 43210, USA}
\newcommand{\aFe}{[$\alpha$/Fe]\xspace}
\newcommand{\vice}{{\tt VICE}\xspace}
\newcommand{\hydro}{{\tt h277}\xspace}
\newcommand{\todo}[1]{{\color{red}#1}}

@@

% Abstract from AAS 241
\begin{abstract}
    Type Ia supernovae (SNe Ia) produce most of the Fe-peak elements in the Universe and therefore are a crucial ingredient in galactic chemical evolution models. SNe Ia do not explode immediately after star formation, and the delay-time distribution (DTD) has not been definitively determined by supernova surveys or theoretical models. Because the DTD also affects the relationship among age, [Fe/H], and \aFe \todo{([O/Fe]?)} in chemical evolution models, comparison with observations of stars in the Milky Way is an important consistency check for any proposed DTD. We implement several popular forms of the DTD in combination with multiple star formation histories for the Milky Way in %one-zone and 
    multi-zone chemical evolution models which include radial stellar migration.
    % using the Versatile Integrator for Chemical Evolution (\vice). \vice includes a prescription for radial stellar migration in multi-zone models. 
    We compare our predicted interstellar medium abundance tracks, stellar abundance distributions, and stellar age distributions to the final data release of the Apache Point Observatory Galactic Evolution Experiment (APOGEE). We find that the DTD has the largest effect on the \aFe distribution: a DTD with more prompt SNe Ia produces a stellar abundance distribution that is skewed toward a lower \aFe ratio.
    % whereas a DTD with more delayed SNe Ia produces a distribution that is skewed toward higher \aFe.
    While the DTD alone cannot explain the observed bimodality in the \aFe distribution, in combination with an appropriate star formation history it affects the goodness of fit between the simulated and observed high-$\alpha$ sequence. Our model results favor an extended DTD with fewer prompt SNe Ia than the fiducial $t^{-1}$ power law.
\end{abstract}

@@

We use \vice to run chemical evolution models which closely follow those of \citet{JohnsonWeinberg2020-Starbursts} and \citet[][hereafter \citetalias{Johnson2021-Migration}]{Johnson2021-Migration}. We refer the interested reader to the former for details about the \vice package and to the latter for details about the model Milky Way disk, including the star formation law, radial density gradient, and outflows. Similar to \citetalias{Johnson2021-Migration}, we adopt a prescription for radial migration based on the \hydro hydrodynamical simulation \citep{Christensen2012-h277}. In Appendix \ref{app:migration}, we describe our method for determining the migration distance $\Delta R_{\rm gal}$ and midplane distance $|z|$ for each model stellar population. Our method produces smoother distributions in chemical abundance space than the simulation-based approach, but the abundance distributions are otherwise unaffected by this change. Table \ref{tab:multizone-parameters} summarizes our model parameters and the sub-sections in which we discuss them in detail.

@@

\paragraph{Early-burst} An extension of the model proposed by \citet{Conroy2022-ThickDisk} to explain the non-monotonic behavior of the high-$\alpha$ sequence down to ${\rm [Fe/H]}\approx-2.5$. This model features an abrupt factor $\sim20$ rise in the SFE at early times, driving an increase in the [O/Fe] abundance at the transition between the epochs of halo and thick disk formation. \citet{Stahlholdt2022-StarFormationEpochs} found evidence for a burst $\sim10$ Gyr ago which marks the beginning of a second phase of star formation. An early infall-driven burst of star formation has been found to lead to MW-like $\alpha$-bimodality in the EAGLE simulations \citep{Mackereth2018-AlphaBimodality} \todo{(cite EAGLE?)}. We adopt the following formula for the time-dependence of the SFE timescale from \citet{Conroy2022-ThickDisk}:
\begin{equation}
    \frac{\tau_{\rm EB}}{1\,\rm{Gyr}} =
    \begin{cases}
        50, & t < 2.5\,\rm{Gyr} \\
        \frac{50}{[1+3(t-2.5)]^2}, & 2.5\leq t \leq 3.7\,\rm{Gyr} \\
        2.36, & t > 3.7\,\rm{Gyr}.
    \end{cases}
    \label{eq:earlyburst-taustar}
\end{equation}
While \citet{Conroy2022-ThickDisk} used a constant infall rate in their one-zone model, we adopt a radially-dependent infall rate which declines exponentially with time:
\begin{equation}
    f_{\rm EB}(t|R_{\rm gal}) = e^{-t/\tau_{\rm sfh}},
    \label{eq:earlyburst-ifr}
\end{equation}
where $\tau_{\rm sfh}$ is the same as in the inside-out case.
To calculate $\dot \Sigma_\star$ from the above quantities, we modify the fiducial star formation law adopted from \citetalias{Johnson2021-Migration}, substituting $\tau_{\rm EB}$ for the SFE timescale of molecular gas:
\begin{equation}
    \dot \Sigma_\star = 
    \begin{cases}
        \Sigma_g \tau_{\rm EB}^{-1}, & \Sigma_g \geq \Sigma_{g,2} \\
        \Sigma_g \tau_{\rm EB}^{-1} \Big(\frac{\Sigma_g}{\Sigma_{g,2}}\Big)^{2.6}, & \Sigma_{g,1} \leq \Sigma_g \leq \Sigma_{g,2} \\
        \Sigma_g \tau_{\rm EB}^{-1} \Big(\frac{\Sigma_{g,1}}{\Sigma_{g,2}}\Big)^{2.6} \Big(\frac{\Sigma_g}{\Sigma_{g,1}}\Big)^{0.7}, & \Sigma_g \leq \Sigma_{g,1},
    \end{cases}
    \label{eq:earlyburst-sfr}
\end{equation}
with $\Sigma_{g,1}=5\times 10^6\,{\rm M}_\odot\,{\rm kpc}^{-2}$ and $\Sigma_{g,2}=2\times 10^7\,{\rm M}_\odot\,{\rm kpc}^{-2}$.

@@

We restrict our sample to red giant branch and red clump stars with high-quality spectra. Table \ref{tab:sample} lists our selection criteria, which largely follow from \citet{Hayden2015-ChemicalCartography}. This produces a final sample of \variable{output/sample_size.txt}stars with [O/Fe] and [Fe/H] abundance measurements. APOGEE stars are matched with their Bailer-Jones photo-geometric distance estimate from \textit{Gaia} Early Data Release 3 \citep{Gaia2016-Mission,Gaia2021-EDR3}, which we use to calculate galactocentric radius $R_{\rm gal}$ and midplane distance $z$.

We use estimated ages from \citet[][hereafter \citetalias{Leung2023-Ages}]{Leung2023-Ages}, who use a variational encoder-decoder network which is trained on asteroseismic data to retrieve age estimates for APOGEE giants without contamination from age-abundance correlations. Importantly, the \citetalias{Leung2023-Ages} ages do not plateau beyond $\sim10$ Gyr as they do in astroNN \citep{Mackereth2019-astroNN-Ages}. We use an age uncertainty cut of 40\% per the recommendations of \citetalias{Leung2023-Ages}, which produces a total sample of \variable{output/age_sample_size.txt}APOGEE stars with age estimates. We note that we use the full sample of \variable{output/sample_size.txt}APOGEE stars unless we explicitly compare to age estimates. Table \ref{tab:sample} also presents the median uncertainties in the data for [Fe/H], [O/Fe], and age.

@@

To quantify the agreement between the MDFs generated by \vice and those observed in APOGEE, we compute the Kullback-Leibler (KL) divergence \citep{KullbackLeibler1951}, defined as
\begin{equation}
    D_{\rm{KL}}(P||Q) \equiv \int_{-\infty}^{\infty} p(x) \log\Big(\frac{p(x)}{q(x)}\Big) dx
    \label{eq:kl-divergence}
\end{equation}
for distributions $P$ and $Q$ with probability density functions (PDFs) $p(x)$ and $q(x)$. If $D_{\rm KL}=0$, the two distributions contain equal information. In this case, $P$ is the APOGEE MDF, $Q$ is the model MDF, and $x={\rm [Fe/H]}$. We bin the distributions with a width of 0.01 dex and apply observational uncertainties to our modeled MDFs according to Table \ref{tab:sample}. For each SFH and DTD, we compute $D_{\rm{KL}}$ in the 18 different Galactocentric regions shown in Figure \ref{fig:feh-df-comparison}. We use bins in $R_{\rm gal}$ with a width of 2 kpc between 3 and 15 kpc, and bins in midplane distance of $|z|=0-0.5$ kpc, $0.5-1$ kpc, and $1-2$ kpc. The score $S$ for the entire model is taken to be the average of $D_{\rm{KL}}$ for each region $R$ weighted by the number of APOGEE stars in that region $N_{\star,R}$:
\begin{equation}
    S = \frac{\sum_R D_{\rm KL}(P||Q|R) N_\star(R)}{\sum_R N_\star(R)}.
    \label{eq:feh-df-score}
\end{equation}

@@

% [Talk about scores]
We again compute the KL divergence (Equation \ref{eq:kl-divergence}) to quantify the agreement between the [O/Fe] DFs of our models and APOGEE. We calculate a score for each model as described in Section \ref{sec:feh-df}. The best-scoring model uses the two-infall SFH with the plateau ($W=0.3$ Gyr) DTD, although both plateau and exponential DTDs as well as the triple-system DTD score well when combined with certain SFHs. The late-burst SFH generally produces [O/Fe] DFs that score poorly compared to other models. \todo{Elaborate on the model scores a bit more}.

@@

% Results for the two-population and power-law ($\alpha=-1.1$) DTDs, with a large fraction of prompt SNe Ia, are presented in the first and second columns, respectively. 
The two-population and power-law ($\alpha=-1.1$) DTDs, which have a large fraction of prompt SNe Ia, produce stellar abundance distributions that are reasonably well-aligned with the APOGEE contours at low $|z|$, but they entirely miss the observed high-$\alpha$ sequence at large $|z|$. The ISM abundance tracks for the 8 kpc zone do not pass through the APOGEE 30\% contour. For both DTDs, the high-[O/Fe] knee is located below the left-most bound of the plot, but we observe a second knee at ${\rm [O/Fe]}\approx0.2$ where the abundance track turns further downward. As discussed in Section \ref{sec:onezone-results}, the second knee is most prominent in the model with the two-population DTD because of its long exponential tail.
% For these models, the simulated stellar abundances are reasonably well-aligned with the APOGEE contours at low $|z|$ (bottom panels), but they entirely miss the observed high-$\alpha$ sequence at high $|z|$ (top panels).

The DTDs with an intermediate median delay time do well at reproducing the observed abundance distributions. The exponential ($\tau=1.5$ Gyr) DTD produces a distribution in Figure \ref{fig:ofe-feh-dtd} which aligns quite well with the outer APOGEE contours in all $|z|$-bins, and even produces a ``ridge'' which extends to high [O/Fe] at low- and mid-latitudes (bottom and center panels, respectively). While it does better at populating the high-$\alpha$ sequence than the previous DTDs, the bulk of the simulated populations at large $|z|$ still fall below the APOGEE 30\% contour.

The two right-hand columns present model results for the plateau ($W=1$ Gyr) and triple-system DTDs, which have the longest median delay times. The high-[O/Fe] knee occurs at a much higher metallicity in these models and is visible in the gas abundance tracks in the upper-left corner of the panels. At large $|z|$, the simulated abundance distributions align quite well with the APOGEE high-$\alpha$ sequence, but there is a significant ridge of high-$\alpha$ stars from the inner Galaxy at low $|z|$.

To quantify the agreement between the multi-zone model outputs and data in [O/Fe]--[Fe/H] space, we implement the method of \citet{PerezCruz2008-KLTest2D} for estimating the KL divergence between two continuous, multivariate samples using a $k$-nearest neighbor estimate. For $n$ samples from a multivariate PDF $p(\mathbf{x})$ and $m$ samples from $q(\mathbf{x})$, we can estimate $D_{\rm KL}(P||Q)$ according to the following:
\begin{equation}
    \hat D_k(P||Q) = \frac dn \sum_{i=1}^n\log\frac{r_k(\mathbf{x}_i)}{s_k(\mathbf{x}_i)} + \log\frac{m}{n-1},
\end{equation}
where $r_k(\mathbf{x}_i)$ and $s_k(\mathbf{x}_i)$ are the distance to the $k$th nearest neighbor of $\mathbf{x}_i$ in the samples of $P$ and $Q$, respectively. We take $k=2$ to find the nearest neighbor other than the sample itself. As before, $P$ is the APOGEE distribution and $Q$ is the model distribution, and in this case $\mathbf{x}=({\rm [Fe/H]}, {\rm [O/Fe]})$.
As in Sections \ref{sec:feh-df} and \ref{sec:ofe-df}, we bin the model outputs and data by $R_{\rm gal}$ and $|z|$, calculate $\hat D_k(P||Q)$ in each region, and then take the weighted mean of each region as in Equation \ref{eq:feh-df-score} to arrive at a single score for each model.
\todo{More detail on the scoring calculations.}
% \footnote{We got the implementation from \url{https://mail.python.org/pipermail/scipy-user/2011-May/029521.html}.}

@@

\todo{I just noticed that I was using slightly wrong yields in all of the multi-zone models (JW20 instead of J+21). I re-ran all the models and this corrected the end-point of the inside-out tracks, which are now at (0,0). However, this also shifted the end-point of the two-infall tracks down to (+0.1, -0.1), so now two-infall is no longer the clear winner like it used to be. The best DTDs are now the 1-Gyr power-law and the triple-system, rather than the 1.5-Gyr exponential like before. The general trend of more prompt SNe = worse does still hold. At a minimum, I think this indicates that these scores are highly sensitive to the yields. At worst, maybe we need to consider adjusting the yields for the two-infall models so that the [O/Fe] distributions are not significantly lower than every other SFH.}

@@

\begin{figure*}
    \centering
    \includegraphics[width=\linewidth]{figures/radial_migration.pdf}
    \caption{The distribution of final radius $R_{\rm final}$ as a function of formation radius $R_{\rm form}$ and age for the \hydro analogue (top row) and Gaussian sampling scheme (bottom row). From left to right, star particles are binned by formation annulus, from the inner disk (far left column) to the solar annulus (center column) to the outer disk (far right column). Within each panel, colored curves represent the different age bins, ranging from the youngest stars (dark blue) to the oldest (dark red). In the top row, we exclude age bins with fewer than 100 unique analogue IDs for clarity. All distributions are normalized so that the area under the curve is 1, and have been smoothed by a boxcar function with a width of 0.5 kpc. The vertical dotted black lines indicate the bounds of each bin in $R_{\rm form}$; stars within that region of the distribution have not migrated outside their birth annulus over their lifetime.}
    \label{fig:radial-migration}
    \script{radial_migration.py}
\end{figure*}

@@

For each star particle in \vice, \citetalias{Johnson2021-Migration} randomly assign an analogue star particle from \hydro and adopt its radial migration distance $\Delta R$ and final midplane distance $z$. This allows \vice to adopt a realistic pattern of radial migration without needing to implement its own hydrodynamical simulation. However, in regions where the number of \hydro star particles is relatively low, such as at large $R_{\rm gal}$ and small $t$, a single \hydro star particle can be assigned as an analogue to multiple \vice stellar populations. These populations will have similar formation and migration histories and consequently similar abundances, which produces unphysical ``clumps'' of stars in the abundance distributions at high latitudes and large radii.

We fit a Gaussian to the distribution of $\Delta R = R_{\rm final} - R_{\rm initial}$ from the \hydro output in bins of $R_{\rm initial}$ and age. Each Gaussian is centered at 0 and we find that the scale $\sigma_{\Delta R}$ is best described by the function
\begin{equation}
    \sigma_{\Delta R} = 1.35\,{\rm kpc} \Big(\frac{R_{\rm form}}{8\,{\rm kpc}}\Big)^{0.61} \Big(\frac{T}{1\,{\rm Gyr}}\Big)^{0.33}
    \label{eq:radial-migration}
\end{equation}
where $T$ is the age of the star particle.

We fit a sech$^2$ function \citep{Spitzer1942} to the distribution of midplane distances $z$. Vertical migration away from the midplane does not affect the chemical evolution model, but we do use $z$ in our analysis. The probability density function (PDF) of $z$ given some scale height $h_z$ is
\begin{equation}
    {\rm PDF}(z) = \frac{1}{4 h_z} {\rm sech}^2\Big(\frac{z}{2 h_z}\Big)
    \label{eq:sech-pdf}
\end{equation}
and the corresponding cumulative distribution function (CDF) is
\begin{equation}
    {\rm CDF}(z) = \frac{1}{1 + e^{-z / h_z}}.
    \label{eq:sech-cdf}
\end{equation}
We fit the above function to the distributions of $z$ in \hydro in varying bins of $T$ and $R_{\rm final}$ and found that $h_z$ is best described by the function
\begin{equation}
    h_z = (0.25\,{\rm kpc}) 
    e^{\frac{T-5\,{\rm Gyr}}{7.0\,{\rm Gyr}}}
    e^{\frac{R_{\rm final}-8\,{\rm kpc}}{6.0\,{\rm kpc}}}.
    \label{eq:scale-height}
\end{equation}

@@

\citet{Okalidis2022-AurigaMigration} studied the stellar migration in the Auriga simulations and found that the migration strength had both a dependence on age and formation radius. They also found that in strongly-barred galaxies, the migration is stronger overall but has a slower timestep evolution than diffusion ($\Delta t^{0.5}$). \todo{They argue that the bar has a minimal effect.}
On the other hand, \citet{Brunetti2011-BarredSpiralDiffusion} argue that the bar has a strong effect on the strength of radial migration and that its effect is not constant with time.

@@

\citet{Lu2023-LMCStars} examined radial migration strengths in galaxies from the NIHAO simulations. They found that it predicts galaxies to follow a relatively universal relation of $\sigma_{\rm RM}\propto \tau^{0.32}$, which is slower than the diffusion process of, e.g., \citet{Frankel2019-InsideOutGrowth}, but quite similar to our own age dependence.
